Enhanced Particle Swarm Optimization In Sketch Recognition

### 2010-08-18 trunk/Swarm/tmp/swarmModified.tex

%where $iRandom_i$ and $Random_j$ are two particles in the solution space. This modification replace the distance between the current particle and the local best with two random particle from the space. The effect of this change in more expolration and avoiding of local min problems. Equation \ref{eq:ModSwarm} is computed and the new particle locaiton is calculated. The new location is compared the location of particle if with equation \ref{eq:Swarm} if the modified location is better (mean better fitness function) then the new location is selected otherwise the old equation is used. 
\section{System Discribtion  }\label{Sysdisc}
%\label{sec:application}
The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{myPaper}. 
%\subsection{The system}

 The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step.
\subsection{Preprocessing}
\label{Prepross}
 Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}. It was noted that time difference provides more distinct maxima than the speed information. Agar \cite{polygonfeedback31} mentions that the pointing device (i.e. the mouse) sampling rate is the reason for this phenomenon. 
 In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$. Dominant points are characterized by lower speed values and higher curvature and direction values.

@@

 where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke. A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.  After few iteration, the final fit error of the estimated ellipse is reported. Another measure is used to compute the efficiency of the final estimated ellipse. Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered. This eliminates fitting a line into very large ellipse but leaves small ellipses to be fitted as a partial or full ellipse.

@@

After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
\begin{enumerate}
	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$. 
	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.  Otherwise, one of these points are removed based on the error cased by its removal.
	\item Each two successive segments are test to determine if they can be merged into a single segment. If merging the two segments have the same or smaller the error, the two segments are merged. This step ensures that the final segmentation of the system has minimum number of segments. 
\end{enumerate}
%1) 2)  % If two points are nearer than $min_D$ then one of the points is removed. 
\subsection{Recognition}
\label{sec:Recognition}
After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process. A composite set of features is used to generate a single feature vector. The features used consist of Rubine feature set,  Zernike moments of order 10 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol. After computing the features the symbol is introduced to the (SVM) classifier.

\cite{myPaper}  
\section{Experiments} 
\label{sec:Experiments}
In our experiment we used the same benchmark dataset used in \cite{myPaper}, the dataset was collected by Hse and Newton\cite{HeloiseBeautification}. The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. Five different splits were generated to divide the dataset into training and test sets. The results displayed are the average recognition accuracy of the five splits. The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.

@@

		\includegraphics[scale=0.5]{symbolSet.PNG} 
		\caption[The Symbol Set] {The symbol set}   \label{fig:symbolSet}
	 \end{figure}
	  %The first experiment with the whole feature set and the second experiments with only half of the features.
To demonstrate the change in accuracy is due to the enhanced swarm optimization algorithm we tested the recognition system using two sets of features. Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) using the whole feature set. The second experiment was tested on both algorithm but only using half the feature set.

Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively. The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.  The increase in recognition rate is consistent in all swarm algorithms and in both experiments.    %n the We performed two experiments to test the system. Firstly, we tested recognition accuracy of shapes in the data set using the orginal swarm segmentation algorithm and then after using the enhanched swarm optimization. To  both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms. Figure \ref{fig:test1} shows the accuracy achieved by each algorithm. The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.

@@

 %The goal of the system is to modify the segmentation and recognition of the sketch system. It made the same type of experiement done in \cite{myPaper}. The dataset is divided into 13 category of goemetric %symbols

% The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.  Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. The datasets were divided into training and test sets. Five different splits for the training and test data are generated from each dataset. The results displayed are the average recognition accuracy of the five splits. The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
    \begin{figure}
  \centering
  \includegraphics[width=3.5in]{Experiment92.jpg}
  \caption[Recognition accuracy]%
  {The recognition rate of system using the feature set.}
   \label{exp2}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=3.5in]{Experiment68.jpg}
  \caption[Recognition accuracy]%
  {The recognition rate of system using half the feature set.}
   \label{exp1}
\end{figure}

>>>>

%where $iRandom_i$ and $Random_j$ are two particles in the solution space. This modification replace the distance between the current particle and the local best with two random particle from the space. The effect of this change in more expolration and avoiding of local min problems. Equation \ref{eq:ModSwarm} is computed and the new particle locaiton is calculated. The new location is compared the location of particle if with equation \ref{eq:Swarm} if the modified location is better (mean better fitness function) then the new location is selected otherwise the old equation is used. 
\section{System Discribtion  }\label{Sysdisc}
%\label{sec:application}
The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{mypaper}. The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step. The main block of the system remains the same but their is some modification in the segmentation algorithm to improve its efficiency. The main modification is using Enhanced Particle swarm algorithm instead of Basic Particle swarm algorithms. Other modifications are explained in the details in the next sections.  
\subsection{Preprocessing}
\label{Prepross}
 Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}. It was noted that time difference provides more distinct maxima than the speed information. Agar \cite{polygonfeedback31} mentions that the pointing device (i.e. the mouse) sampling rate is the reason for this phenomenon. 
 In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$. Dominant points are characterized by lower speed values and higher curvature and direction values.

@@

 where $N$ is number of points in the stroke, $a,b$ are the length of ellipse axes, $x_0$ \& $y_0$ are the coordinates of the center point, $x_i$ \& $y_i$ are the coordinates of point $i$ in the stroke. A list of new values for $x_0$ , $y_0$ ,$a$ and $b$ are generated randomly from the older values with small increments after each loop.  After few iteration, the final fit error of the estimated ellipse is reported. A modification to the system in \cite{mypaper} is trying to add another measure to compute the efficiency of the final estimated ellipse. Equation \ref{eq:circleError} ensures that the drawn percentage of ellipse is considered. This eliminates fitting a line into very large ellipse but leaves small ellipses to be fitted as a partial or full ellipse.

@@

As a modification to the system in \cite{mypaper}. After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
\begin{enumerate}
	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$. 
	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.  Otherwise, one of these points are removed based on the error cased by its removal.
	\item Each two successive segments are test to determine if they can be merged into a single segment. If merging the two segments have the same or smaller the error, the two segments are merged. This step ensures that the final segmentation of the system has minimum number of segments. 
\end{enumerate}
%1) 2)  % If two points are nearer than $min_D$ then one of the points is removed. 
\subsection{Recognition}
\label{sec:Recognition}
After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process. A composite set of features is used to generate a single feature vector. The features used consist of Rubine feature set,  Zernike moments of order 10 \cite{HeloiseBeautification}, ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol. After computing the features the symbol is introduced to the (SVM) classifier.

\cite{mypaper}  
\section{Experiments} 
\label{sec:Experiments}
In our experiment we used the same benchmark dataset used in \cite{mypaper}, the dataset was collected by Hse and Newton\cite{HeloiseBeautification}. The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. Five different splits were generated to divide the dataset into training and test sets. The results displayed are the average recognition accuracy of the five splits. The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.

@@

		\includegraphics[scale=0.5]{symbolSet.PNG} 
		\caption[The Symbol Set] {The symbol set}   \label{fig:symbolSet}
	 \end{figure}
	  %The first experiment with the whole feature set and the second experiments with only half of the features.
To demonstrate the change in accuracy is due to the enhanced swarm optimization algorithm we tested the recognition system using two sets of features. Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) using the whole feature set. The second experiment was tested on both algorithm but only using half the feature set. Both experiments were compared with the original sketch recognition system presented in \cite{mypaper}.

Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively. The figures compares between the result that can be obtained by system in \cite{mypaper} and modifying the system using Enhanced PSO. The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.  The increase in recognition rate is consistent in all swarm algorithms and in both experiments.   %n the We performed two experiments to test the system. Firstly, we tested recognition accuracy of shapes in the data set using the orginal swarm segmentation algorithm and then after using the enhanched swarm optimization. To  both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms. Figure \ref{fig:test1} shows the accuracy achieved by each algorithm. The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1.

@@

 %The goal of the system is to modify the segmentation and recognition of the sketch system. It made the same type of experiement done in \cite{mypaper}. The dataset is divided into 13 category of goemetric %symbols

% The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system.  Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. The datasets were divided into training and test sets. Five different splits for the training and test data are generated from each dataset. The results displayed are the average recognition accuracy of the five splits. The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test.
    \begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment92.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of the full feature set on both the original system presented in \cite{mypaper} and Enhanced PSO. The results shows that Enhanced PSO increases the performance over the original system. }
   \label{exp2}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment68.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of half the feature set on both the original system presented in \cite{mypaper} and Enhanced PSO. The results shows that Enhanced PSO increase the recognition rate. }
   \label{exp1}
\end{figure}
